# 110 Metabolomics RNAseq

小鼠血液和肝脏的代谢组，和肝脏的转录组数据分析。转录组数据只有两个生物重复。

```{r, include=F, echo=F, message=F, warnings=F}
options(java.parameters = '-Xmx8000m')
source('Rscript/setup.R')
source('Rscript/functions.R')
source('Rscript/app_functions.R')
source('rscript/pathway_funs.R', encoding = 'UTF-8')


options(useFancyQuotes = FALSE)


knitr::opts_chunk$set(
  cache = T,
  message = F,
  dpi = 300
)

library(plyr)
library(dplyr)
library(ggplot2)
library(magrittr)
library(kableExtra)
library(pathview)

library(extrafont)
loadfonts(device = 'win', quiet = T)
par(family = 'Liberation')
windowsFonts(
  serif = 'Liberation Serif',
  sans = 'Liberation Sans',
  mono = 'Liberation Mono'
)

par(family = 'sans')

set.seed(1)

prefix <- '110'
output.dir <- file.path('output', prefix)
Sys.setlocale("LC_ALL","Chinese")

if (!dir.exists(output.dir)) {
    dir.create(output.dir, recursive = T)
}
```

```{r}
d.liver.rna <- read.csv(
    'data/肝脏转录组比较_for代谢分析.csv',
    check.names = F
)

d10 <- d.liver.rna
```

## Setup MetaboAnalystR

```{r}
## this variable is used by .get.my.lib functions in pathway_funs.R
metaboanalyst.db.path <- 'database/metaboanalyst/20210113'

## load file from pre-defined directory
## latest database/metaboanalyst/20210113
## rda files from database/metaboanalyst/MetaboAnalyst4.93/
environment(get.my.lib) <- asNamespace('MetaboAnalystR')
assignInNamespace('.get.my.lib', get.my.lib, ns = 'MetaboAnalystR')

## load the same database as in MetaboliteMappingExact and CreateMappingResultTable 
environment(GetFinalNameMap2) <- asNamespace('MetaboAnalystR')
assignInNamespace('GetFinalNameMap', GetFinalNameMap2, ns = 'MetaboAnalystR')

## able to load qs or rda file from local directory
environment(SetCurrentMsetLib2) <- asNamespace('MetaboAnalystR')
assignInNamespace('SetCurrentMsetLib', SetCurrentMsetLib2, ns = 'MetaboAnalystR')

## able to load rda files from database/metaboanalyst/MetaboAnalyst4.93/
environment(CalculateHyperScore2) <- asNamespace('MetaboAnalystR')
assignInNamespace('CalculateHyperScore', CalculateHyperScore2, ns = 'MetaboAnalystR')


environment(PlotORA2) <- asNamespace('MetaboAnalystR')
assignInNamespace('PlotORA', PlotORA2, ns = 'MetaboAnalystR')

environment(PlotEnrichDotPlot2) <- asNamespace('MetaboAnalystR')
assignInNamespace('PlotEnrichDotPlot', PlotEnrichDotPlot2, ns = 'MetaboAnalystR')

## adjust margin
environment(PlotMSEA.Overview2) <- asNamespace('MetaboAnalystR')
assignInNamespace('PlotMSEA.Overview', PlotMSEA.Overview2, ns = 'MetaboAnalystR')

## able to load qs or rda file from local directory
environment(SetKEGG.PathLib2) <- asNamespace('MetaboAnalystR')
assignInNamespace('SetKEGG.PathLib', SetKEGG.PathLib2, ns = 'MetaboAnalystR')

## able to load rda files from database/metaboanalyst/MetaboAnalyst4.93/
environment(CalculateOraScore2) <- asNamespace('MetaboAnalystR')
assignInNamespace('CalculateOraScore', CalculateOraScore2, ns = 'MetaboAnalystR')

## make sure CrossReferencing and CreateMappingResultTable load the same databases
environment(CreateMappingResultTable2) <- asNamespace('MetaboAnalystR')
assignInNamespace('CreateMappingResultTable', CreateMappingResultTable2, ns = 'MetaboAnalystR')


## must load after assignInNamespace
library(MetaboAnalystR)
```

## Liver RNA seq

4.1% of input gene IDs are fail to map

```{r}
library(AnnotationDbi)
library(clusterProfiler)
library(topGO)
library(rlist)
## http://bioconductor.org/packages/3.12/data/annotation/
library(org.Mm.eg.db)

############
# 加载基因、orgdb
############
message("load gene/orgdb")
gene     <- d10$ID
org_db <- org.Mm.eg.db

message("convert SYMBOL to ENTREZID")
eg   <- bitr(gene, fromType="SYMBOL", toType="ENTREZID", OrgDb=org_db)

d10 <- dplyr::left_join(d10, eg, by = c('ID' = 'SYMBOL'))

d10.entrez <- d10 %>%
    filter(
        !is.na(ENTREZID)
    ) %>%
    tibble::column_to_rownames('ENTREZID')


# AnnotationDbi::mget(gene, org.Mm.egSYMBOL, ifnotfound=NA)

```

```{r, eval=F}
############
# GO富集分析
############
message("start GO analysis")
message('CC')
ego_cc <- enrichGO(gene          = gene,
                   OrgDb         = org_db,
                   ont           = "CC",
                   pAdjustMethod = "BH",
                   pvalueCutoff  = 0.05,
                   qvalueCutoff  = 1,
                   readable      = TRUE)
message('BP')
ego_bp <- enrichGO(gene          = gene,
                   OrgDb         = org_db,
                   ont           = "BP",
                   pAdjustMethod = "BH",
                   pvalueCutoff  = 0.05,
                   qvalueCutoff  = 1,
                   readable      = TRUE)
message('MF')
ego_mf <- enrichGO(gene          = gene,
                   OrgDb         = org_db,
                   ont           = "MF",
                   pAdjustMethod = "BH",
                   pvalueCutoff  = 0.05,
                   qvalueCutoff  = 1,
                   readable      = TRUE)

############
# BP/CC/MF 绘图
############
go.pdf <- file.path(output.dir, "enrich_go.pdf")
                       
cairo_pdf(go.pdf, onefile = T, w = 12, h = 5, family = 'sans')
dotplot(ego_bp,showCategory=20)
dotplot(ego_cc,showCategory=20)
dotplot(ego_mf,showCategory=20)
dev.off()

go.xlsx <- file.path(output.dir, 'enrich_go.xlsx')

wb <- XLConnect::loadWorkbook(go.xlsx, create = T)
XLConnect::createSheet(wb, 'bp')
XLConnect::writeWorksheet(wb, ego_bp, 'bp')

XLConnect::createSheet(wb, 'cc')
XLConnect::writeWorksheet(wb, ego_cc, 'cc')

XLConnect::createSheet(wb, 'mf')
XLConnect::writeWorksheet(wb, ego_mf, 'mf')

XLConnect::saveWorkbook(wb)
XLConnect::xlcFreeMemory()

############
# KEGG分析
############
message("start KEGG analysis")

file.kegg_enrichment <- file.path(output.dir, "enrichment_kegg.xls")

kk <- enrichKEGG(gene         = gene,
                 organism     = 'mmu',
                 keyType = 'ncbi-geneid',
                 pvalueCutoff = 0.05,
                 qvalueCutoff = 1)

# kk_enrich <- as.data.frame(kk)
# kk_enrich$geneID <- unlist(lapply(1:length(kk_enrich$geneID), function(t){  paste(eg[[1]][eg[[2]] %in% unlist(strsplit(kk_enrich$geneID[t], "/"))], collapse = "/")  }))

kegg_pdf <- file.path(output.dir, "enrichment_kegg.pdf")
if(nrow(kk_enrich) >= 1){
	cairo_pdf(kegg_pdf, onefile = T, height=5,width=12, family = 'sans')
	dotplot(kk,showCategory=20) %>% print
	dev.off()
	XLConnect::writeWorksheetToFile(
	  file.kegg_enrichment, kk_enrich, sheet = 'KEGG', rownames=F)
}



```


## Pathview

```{r}
##------------------------------------------------------
## Pathview
##------------------------------------------------------
pathviewWrap <- function(gene.ht, cpd.ht, pathways, organism,
                         out.suffix, kegg.dir, out.dir.1, out.dir.2) {
    ## kegg.dir is relative path from out.dir
    ## out.dir.1 is the relative path to store pathview map
    ## out.dir.2 is the relative path to store annotated pathway map
    ## pathview-generated pathway maps are saved in current directory
    if (!dir.exists(kegg.dir))
        dir.create(kegg.dir, recursive = T)
    
    if (!dir.exists(out.dir.1))
        dir.create(out.dir.1, recursive = T)
    
    if (!dir.exists(out.dir.2))
        dir.create(out.dir.2, recursive = T)
    
    gene.data <- structure(gene.ht[, 'log2FoldChange'],
                           names = gene.ht[, 'entrez'])
    
    gene.ht %<>% 
        dplyr::mutate(
            ## label shown on the image
            label_img = paste0(
                ID,
                ifelse(padj < 0.05,
                       ifelse(log2FoldChange < 0, '-', '+'),
                       '')
            )
        )
    
    cpd.data <- cpd.ht %>%
        filter(KEGG != '') %>%
        (function(x) {structure(x$log2FC, names = x$KEGG)})

    ## names(table) remove NA
    plyr::l_ply(
        pathways, 
        function(pw.i) {
            ## original    
            file.png <- paste(pw.i, out.suffix, 'png', sep = '.')
            ## pathview
            img.file <- file.path(out.dir.1, file.png)
            ## pathview annotate
            out.file <- file.path(out.dir.2, 
                                  stringi::stri_replace_all_fixed(file.png, 'png', 'pdf'))

            if (file.exists(img.file) & file.exists(out.file)) {
                print(paste('Skip', img.file))
                return(invisible())
            }
            
            tryCatch(
                pv.out <- pathview::pathview(
                    gene.data = gene.data,
                    cpd.data = cpd.data,
                    pathway.id = pw.i, specie = organism, 
                    out.suffix = out.suffix,
                    keys.align = "y", kegg.native = T, 
                    kegg.dir = kegg.dir,
                    key.pos = 'topright',
                    node.sum = 'max.abs'
                ),
                error = function(e) {print(e); print('pw.i')}
            )

            ## save plot.data.gene and plot.data.cpd
            ## show only top 3 genes
            plot.data.gene <- pv.out$plot.data.gene %>%
                dplyr::left_join(gene.ht, by = c('kegg.names' = 'entrez')) %>%
                dplyr::rename(entrez = kegg.names) %>%
                dplyr::mutate(
                    label_img = (function(x) {
                        entrez.ids <- stringi::stri_split_fixed(x, ',') %>%
                            purrr::map_chr((function(y) {
                                gene.ht %>%
                                    dplyr::filter(entrez %in% y) %>%
                                    dplyr::arrange(desc(abs(log2FoldChange))) %>%
                                    select(label_img) %>%
                                    unlist %>% 
                                    (function(x) {
                                        if (length(x) > 3) {
                                            x <- x[1:4]
                                            x[4] <- '...'
                                        }
                                        return(paste0(x, collapse = ','))
                                    })()
                            }))
                    })(all.mapped)
                ) %>%
                dplyr::filter(
                    !is.na(ID)
                )
            
            if (nrow(plot.data.gene) > 0) {
                file.gene <- file.path(out.dir.2, paste(pw.i, out.suffix, 'gene.csv', sep = '.'))
                if (!file.exists(file.gene)) write.csv(plot.data.gene, file.gene)
            }
            
            plot.data.cpd <- pv.out$plot.data.cpd %>%
                dplyr::rename(KEGG = kegg.names) %>%
                dplyr::left_join(cpd.ht, by = c('KEGG')) %>%
                dplyr::mutate(
                    label = ifelse(`non-parametric pvalue` < 0.05, '*', '')
                ) %>%
                dplyr::filter(
                    !is.na(Metabolite)
                )
            
            if (nrow(plot.data.cpd) > 0) {
                file.cpd <- file.path(out.dir.2, paste(pw.i, out.suffix, 'compound.csv', sep = '.'))
                if (!file.exists(file.cpd)) write.csv(plot.data.cpd, file.cpd)
            }
            

            
            
            ## move pathway png file to out.dir.1
            if (file.exists(file.png)) {
                
                if (file.rename(file.png, img.file)) {
                    
                    if (nrow(plot.data.gene) == 0 & nrow(plot.data.cpd) == 0) {
                        return(invisible())
                    }
                    
                    ## read png file
                    img <- png::readPNG(img.file)
                    
                    ## output image parameters
                    width <- ncol(img)
                    height <- nrow(img)
                    res <- 100
                    
                    ## print the original image first
                    # tiff(out.file, width = width, height = height, res = res, compression = 'lzw+p')
                    pdf(out.file, w = width / res, h = height / res)
                    par(mar = c(0, 0, 0, 0))
                    plot(c(0, width), c(0, height), type = "n", xlab = "", 
                         ylab = "", xaxs = "i", yaxs = "i")
                    rasterImage(img, 0, 0, width, height, interpolate = F)
                    
                    if (nrow(plot.data.gene) > 0) {
                        text(x = plot.data.gene$x, 
                             y = height - plot.data.gene$y + plot.data.gene$height - 2, 
                             labels = plot.data.gene$label_img, col = 'blue', 
                             cex = 0.8)
                    }
                    
                    if (nrow(plot.data.cpd) > 0) {
                        text(x = plot.data.cpd$x + plot.data.cpd$width, 
                             y = height - plot.data.cpd$y + plot.data.cpd$height,
                             labels = plot.data.cpd$label, col = 'red', cex = 0.8)
                    }
                    
                    dev.off()
                }
            }
        })
    
    
    # stopCluster(cl)
    invisible()
}

```

## Blood metabolite

```{r, eval=F}
d10.blood <- XLConnect::readWorksheetFromFile(
    'data/20201118/2020-006-C-01-I blood statistical format.xlsx',
    sheet = 'data.csv',
    rownames = 1,
    check.names = F
)[1:12, ]

s10.blood <- XLConnect::readWorksheetFromFile(
    'data/20201118/2020-006-C-01-I blood statistical format.xlsx',
    sheet = 'sample.csv',
    rownames = 1,
    check.names = F
) %>%
    mutate(
        Group = factor(Group, levels = c('WT', 'A'))
    )

v10.blood <- XLConnect::readWorksheetFromFile(
    'data/20201118/2020-006-C-01-I blood statistical format.xlsx',
    sheet = 'var.csv',
    rownames = 1,
    check.names = F
) %>%
    filter(
        `Class total` != 'Class'
    )


## HMDB ID version 4
d.var <- (v10.blood %>% dplyr::mutate(
  `Metabolite Name` = rownames(v10.blood),
  `HMDB ID` = stringi::stri_extract_first_regex(`Database ID`, 'HMDB[0-9]{7}')
))[, c('Metabolite Name', 'Class', 'Database ID', 'HMDB ID')]


## load HMDB database
hmdb <- read.csv(file.path('../../template/', 'database/HMDB/20190514/hmdb_metabolites.csv'), check.names = F, row.names = 1)
colnames(hmdb) <- c('HMDB ID', 'HMDB_name', 'HMDB_direct_parent', 'HMDB_class')

d.var <- dplyr::left_join(d.var, hmdb, by = 'HMDB ID')

## retrieve KEGG ID
tmp <- capture.output(
  id <- idMap(d.var[, 'Database ID'])
)

testthat::expect(
  !any(duplicated(id[!(id[, 'HMDB'] %in% c('', 'NA')), 'HMDB'])),
  failure_message = 'Duplicated HMDB ID'
)

d.var <- dplyr::left_join(d.var, data.frame(id), by = c('Database ID' = 'Query'))
# d.var <- dplyr::left_join(d.var, data.frame(id), by = c('HMDB_v4' = 'HMDB'))

d.var <- lapply(d.var, function(x) {
  ifelse(is.na(x), '', x)
}) %>% data.frame(check.names = F)

ht.blood <- hypothesisTest(d10.blood, as.factor(s10.blood$Group))

ht.blood %<>% 
    as.data.frame() %>%
    mutate(
        log2FC = log2(`Fold: A/WT`)
    )

testthat::expect_equal(
    rownames(ht.blood),
    d.var[, 'Metabolite Name']
)
# 
# ht.blood %>%
#     mutate(
#         KEGG = d.var$KEGG,
#         HMDB = d.var$`HMDB ID`
#     ) %>%
#     filter(
#         T
#     ) %>%
#     select(
#         everything()
#     ) %>%
#     write.table(
#         'compound_blood.txt', row.names = T, sep = '\t', quote = F
#     )

ht.blood %<>%
    mutate(
        KEGG = d.var$KEGG,
        kegg.names = '',
        HMDB = d.var$`HMDB ID`
    )

for (i in seq_along(ht.blood$kegg.names)) {
    if (ht.blood[i, 'KEGG'] != '') {
        tryCatch(
            ht.blood[i, 'kegg.names'] <- KEGGREST::keggFind('compound', ht.blood[i, 'KEGG']),
            error = function(e) {print(e)}
        )
    }
}

local({
    f <- 'output/110/hypothesis_test_blood.csv'
    if (!file.exists(f)) {
        write.csv(ht.blood, f)
    }
})
```

```{r, eval=F}
cmpd.vec <- unique(ht.blood$KEGG) %>%
    Filter(function(x) x != '', .)

cmpd.vec[45] <- 'C04741'

plyr::l_ply(cmpd.vec, function(x) {cat(x, '\n')})
```

```{r, eval=F}
mSet<-InitDataObjects("conc", "pathora", FALSE)
# cmpd.vec <- c("C00349","C00157","C09908","C02530","C00846","C17687","C00109","C00399","C00136","C01190","C08788","C17230","C01033","C07632","C04741","C14441","C07185","C04554","C07635","C14831","C05849","C02504","C00535","C06423","C12989","C05455","C00376","C00427","C06439","C00334","C02939","C00356","C02593","C01949","C05565","C00858","C15610","C02744","C04737","C16218","C00882","C08363","C16678","C00410","C15786","C04272","C04411","C06573","C01907","C07369","C03205","C01089","C14809","C02360","C08278","C15915","C01933","C03665","C06255","C09727","C01780","C02406","C04373","C06135","C05672","C17647","C14717","C01164","C17438","C01571","C17664","C01953","C02528","C04661","C00249")
mSet<-Setup.MapData(mSet, cmpd.vec);
mSet<-CrossReferencing(mSet, "kegg");
mSet<-CreateMappingResultTable(mSet)
# mSet<-PerformDetailMatch(mSet, "C12989");
# mSet<-GetCandidateList(mSet);
mSet<-SetKEGG.PathLib(mSet, "mmu", "current")
mSet<-SetMetabolomeFilter(mSet, F);
mSet<-CalculateOraScore(mSet, "rbc", "hyperg")
```


### Metabolic pathways

Obtained metabolic pathways hit by the metabolites using MetPA.

```{r}
load('RData/blood.RData')

pathways <- KEGGREST::keggList('pathway', 'mmu')

d.pathways.blood <- purrr::imap(pathways, function(x, y) {
    c(substr(y, 6, 13),
      stringi::stri_replace_all_fixed(x, ' - Mus musculus (mouse)', '')
    )
}) %>% 
    do.call(rbind, .) %>%
    as.data.frame %>%
    magrittr::set_colnames(c('code', 'pathway_name')) %>%
    dplyr::right_join(
        mSet$analSet$ora.mat %>% tibble::rownames_to_column('pathway_name'),
        by = 'pathway_name'
    )

if (!dir.exists('RData')) {
    dir.create('RData')
}
if (!file.exists('RData/blood.RData')) {
    save(ht.blood, d.pathways.blood, mSet, file = 'RData/blood.RData')
}
    

local({
    f <- 'output/110/blood_metabolites_pathways.csv'
    if (!file.exists(f)) {
        write.csv(d.pathways.blood, f)
    }
})

```

```{r}
library(pathview)

cpd.ht <- ht.blood %>%
    tibble::rownames_to_column('Metabolite')

gene.ht <- d10.entrez %>%
    tibble::rownames_to_column('entrez')

pathways.blood <- structure(d.pathways.blood$code,
                            names = d.pathways.blood$pathway_name)

pathviewWrap(gene.ht, cpd.ht, pathways = pathways.blood, organism = 'mmu',
             out.suffix = 'blood', 
             kegg.dir = 'output/110/kegg',
             out.dir.1 = 'output/110/pathview_blood',
             out.dir.2 = 'output/110/pathview_blood_annotate')

gene.summary <- lapply(
    list.files('output/110/pathview_blood_annotate', '*.gene.csv', full.names = T),
    function(f) {
        d <- read.csv(f)
        
        pathway.code <- fs::path_file(f) %>% 
            stringi::stri_split_fixed('.', simplify = T) %>%
            `[`(1,1)
        
        gene.ht %>%
            dplyr::filter(
                entrez %in% 
                    (stringi::stri_split_fixed(d$all.mapped, ',') %>%
                    do.call(c, .) %>%
                    unique)
            ) %>% 
            dplyr::group_by(
                direction = ifelse(
                    padj < 0.05,
                    ifelse(log2FoldChange < 0, 'down', 'up'),
                    'nonsig')
            ) %>%
            dplyr::summarise(
                n = n(),
                names = paste(sprintf('%s(%s)', ID, entrez), collapse = ','),
                .groups = 'drop'
            ) %>%
            dplyr::mutate(
                pathway = pathway.code,
                type = 'gene'
            ) %>%
            dplyr::select(
                pathway, type, direction, n, names
            )
    }) %>%
    do.call(rbind, .)

cpd.summary <- lapply(
    list.files('output/110/pathview_blood_annotate', '*.compound.csv', full.names = T),
    function(f) {
        d <- read.csv(f, check.names = F, row.names = 1)
        
        pathway.code <- fs::path_file(f) %>% 
            stringi::stri_split_fixed('.', simplify = T) %>%
            `[`(1,1)
        
        d %>%
            dplyr::group_by(
                direction = ifelse(
                    `non-parametric pvalue` < 0.05,
                    ifelse(`Fold: A/WT` < 1, 'down', 'up'),
                    'nonsig')
            ) %>%
            dplyr::summarise(
                n = n(),
                names = paste(sprintf('%s(%s; %s)', Metabolite, KEGG, kegg.names), collapse = ','),
                .groups = 'drop'
            ) %>%
            dplyr::mutate(
                pathway = pathway.code,
                type = 'compound'
            ) %>%
            dplyr::select(
                pathway, type, direction, n, names
            )
    }) %>%
    do.call(rbind, .)

gene.cpd.summary <- rbind(gene.summary, cpd.summary) %>%
    as.data.frame %>%
    dplyr::arrange(pathway, type, direction)

local({
    f <- 'output/110/blood_pathway_summary.xlsx'
    if (!file.exists(f)) {
        XLConnect::writeWorksheetToFile(f, gene.cpd.summary, sheet = 'summary')
    }
})


```

## Liver metabolite

```{r}
d10.liver <- XLConnect::readWorksheetFromFile(
    'data/20201118/2020-006-C-01-I liver statistical format.xlsx',
    sheet = 'data.csv',
    rownames = 1,
    check.names = F
)[1:16, ]

s10.liver <- XLConnect::readWorksheetFromFile(
    'data/20201118/2020-006-C-01-I liver statistical format.xlsx',
    sheet = 'sample.csv',
    rownames = 1,
    check.names = F
) %>%
    mutate(
        Group = factor(Group, levels = c('WT', 'A'))
    )


v10.liver <- XLConnect::readWorksheetFromFile(
    'data/20201118/2020-006-C-01-I liver statistical format.xlsx',
    sheet = 'var.csv',
    rownames = 1,
    check.names = F
) %>%
    filter(
        `Class total` != 'Class'
    )


## HMDB ID version 4
d.var <- (v10.liver %>% dplyr::mutate(
  `Metabolite Name` = rownames(v10.liver),
  `HMDB ID` = stringi::stri_extract_first_regex(`Database ID`, 'HMDB[0-9]{7}')
))[, c('Metabolite Name', 'Class', 'Database ID', 'HMDB ID')]


## load HMDB database
hmdb <- read.csv(file.path('../../template/', 'database/HMDB/20190514/hmdb_metabolites.csv'), check.names = F, row.names = 1)

# hmdb <- read.csv(file.path('../../template/', hmdb_db$file), check.names = F, row.names = 1)
colnames(hmdb) <- c('HMDB ID', 'HMDB_name', 'HMDB_direct_parent', 'HMDB_class')

d.var <- dplyr::left_join(d.var, hmdb, by = 'HMDB ID')

## retrieve KEGG ID
tmp <- capture.output(
  id <- idMap(d.var[, 'Database ID'])
)

testthat::expect(
  !any(duplicated(id[!(id[, 'HMDB'] %in% c('', 'NA')), 'HMDB'])),
  failure_message = 'Duplicated HMDB ID'
)

d.var <- dplyr::left_join(d.var, data.frame(id), by = c('Database ID' = 'Query'))
# d.var <- dplyr::left_join(d.var, data.frame(id), by = c('HMDB_v4' = 'HMDB'))

d.var <- lapply(d.var, function(x) {
  ifelse(is.na(x), '', x)
}) %>% data.frame(check.names = F)

ht.liver <- hypothesisTest(d10.liver, as.factor(s10.liver$Group))

ht.liver %<>% 
    as.data.frame() %>%
    mutate(
        log2FC = log2(`Fold: A/WT`)
    )

testthat::expect_equal(
    rownames(ht.liver),
    d.var[, 'Metabolite Name']
)

ht.liver %<>%
    mutate(
        KEGG = d.var$KEGG,
        kegg.names = '',
        HMDB = d.var$`HMDB ID`
    )

for (i in seq_along(ht.liver$kegg.names)) {
    if (ht.liver[i, 'KEGG'] != '') {
        tryCatch(
            ht.liver[i, 'kegg.names'] <- KEGGREST::keggFind('compound', ht.liver[i, 'KEGG']),
            error = function(e) {print(e)}
        )
    }
}


local({
    f <- 'output/110/hypothesis_test_liver.csv'
    if (!file.exists(f)) {
        write.csv(ht.liver, f)
    }
})
```


```{r}
cmpd.vec <- unique(ht.liver$KEGG) %>%
    Filter(function(x) x != '', .)

plyr::l_ply(cmpd.vec, function(x) {cat(paste0('"', x, '", '))})
```

```{r}
mSet<-InitDataObjects("conc", "pathora", FALSE)
cmpd.vec <- c("C02862", "C02571", "C02838", "C02990", "C03017", "C07471", "C00864", "C00836", "C01152", "C01157", "C00643", "C00719", "C00327", "C00037", "C00651", "C00041", "C00062", "C00152", "C00049", "C00064", "C00135", "C00407", "C00123", "C00047", "C00073", "C00079", "C00188", "C00078", "C00082", "C00183", "C02989", "C00437", "C03793", "C01042", "C03145", "C00077", "C00127", "C00408", "C10172", "C01879", "C00213", "C03626", "C00245", "C01179", "C21298", "C00108", "C00587", "C01586", "C00695", "C01921", "C05465", "C05122", "C15497", "C02781", "C00878", "C08349", "C00191", "C00257", "C19910", "C00158", "C00122", "C00489", "C00311", "C00042", "C20827", "C06427", "C06429", "C02678", "C01595", "C00712", "C00249", "C08362", "C14762", "C04230", "C01233", "C01089", "C05984", "C02630", "C01991", "C03547", "C00256", "C00149", "C01551", "C08493", "C05635", "C02043", "C00109", "C00233", "C00322", "C00026", "C00022", "C07338", "C01367", "C02067", "C00346", "C02180", "C00328", "C00633", "C05607", "C03672", "C01120", "C06124", "C02494", "C04545", "C00294", "C00144", "C00130", "C04152", "C00262", "C07599", "C00366", "C00385", "C00847", "C05842", "C00153", "C00299", "C00043", "C01762", "C00178", "C00106", "C01996", "C00114", "C00318", "C11135")
mSet<-Setup.MapData(mSet, cmpd.vec);
mSet<-CrossReferencing(mSet, "kegg");
mSet<-CreateMappingResultTable(mSet)
# mSet<-PerformDetailMatch(mSet, "C12989");
# mSet<-GetCandidateList(mSet);
mSet<-SetKEGG.PathLib(mSet, "mmu", "current")
mSet<-SetMetabolomeFilter(mSet, F);
mSet<-CalculateOraScore(mSet, "rbc", "hyperg")
```

node.sum = 'max.abs'

### Metabolic pathways

Obtained metabolic pathways hit by the metabolites using MetPA.

```{r}
load('RData/liver.RData')

pathways <- KEGGREST::keggList('pathway', 'mmu')

d.pathways.liver <- purrr::imap(pathways, function(x, y) {
    c(substr(y, 6, 13),
      stringi::stri_replace_all_fixed(x, ' - Mus musculus (mouse)', '')
    )
}) %>% 
    do.call(rbind, .) %>%
    as.data.frame %>%
    magrittr::set_colnames(c('code', 'pathway_name')) %>%
    dplyr::right_join(
        mSet$analSet$ora.mat %>% tibble::rownames_to_column('pathway_name'),
        by = 'pathway_name'
    )

if (!dir.exists('RData')) {
    dir.create('RData')
}
if (!file.exists('RData/liver.RData')) {
    save(ht.liver, d.pathways.liver, mSet, file = 'RData/liver.RData')
}
    

local({
    f <- 'output/110/liver_metabolites_pathways.csv'
    if (!file.exists(f)) {
        write.csv(d.pathways.liver, f)
    }
})

```

```{r}
library(pathview)

cpd.ht <- ht.liver %>%
    tibble::rownames_to_column('Metabolite')

gene.ht <- d10.entrez %>%
    tibble::rownames_to_column('entrez')

pathways.liver <- structure(d.pathways.liver$code,
                            names = d.pathways.liver$pathway_name)

pathviewWrap(gene.ht, cpd.ht, pathways = pathways.liver, organism = 'mmu',
             out.suffix = 'liver', 
             kegg.dir = 'output/110/kegg',
             out.dir.1 = 'output/110/pathview_liver',
             out.dir.2 = 'output/110/pathview_liver_annotate')

gene.summary <- lapply(
    list.files('output/110/pathview_liver_annotate', '*.gene.csv', full.names = T),
    function(f) {
        d <- read.csv(f)
        
        pathway.code <- fs::path_file(f) %>% 
            stringi::stri_split_fixed('.', simplify = T) %>%
            `[`(1,1)
        
        gene.ht %>%
            dplyr::filter(
                entrez %in% 
                    (stringi::stri_split_fixed(d$all.mapped, ',') %>%
                    do.call(c, .) %>%
                    unique)
            ) %>% 
            dplyr::group_by(
                direction = ifelse(
                    padj < 0.05,
                    ifelse(log2FoldChange < 0, 'down', 'up'),
                    'nonsig')
            ) %>%
            dplyr::summarise(
                n = n(),
                names = paste(sprintf('%s(%s)', ID, entrez), collapse = ','),
                .groups = 'drop'
            ) %>%
            dplyr::mutate(
                pathway = pathway.code,
                type = 'gene'
            ) %>%
            dplyr::select(
                pathway, type, direction, n, names
            )
    }) %>%
    do.call(rbind, .)

cpd.summary <- lapply(
    list.files('output/110/pathview_liver_annotate', '*.compound.csv', full.names = T),
    function(f) {
        d <- read.csv(f, check.names = F, row.names = 1)
        
        pathway.code <- fs::path_file(f) %>% 
            stringi::stri_split_fixed('.', simplify = T) %>%
            `[`(1,1)
        
        d %>%
            dplyr::group_by(
                direction = ifelse(
                    `non-parametric pvalue` < 0.05,
                    ifelse(`Fold: A/WT` < 1, 'down', 'up'),
                    'nonsig')
            ) %>%
            dplyr::summarise(
                n = n(),
                names = paste(sprintf('%s(%s; %s)', Metabolite, KEGG, kegg.names), collapse = ','),
                .groups = 'drop'
            ) %>%
            dplyr::mutate(
                pathway = pathway.code,
                type = 'compound'
            ) %>%
            dplyr::select(
                pathway, type, direction, n, names
            )
    }) %>%
    do.call(rbind, .)

gene.cpd.summary <- rbind(gene.summary, cpd.summary) %>%
    as.data.frame %>%
    dplyr::arrange(pathway, type, direction)

local({
    f <- 'output/110/liver_pathway_summary.xlsx'
    if (!file.exists(f)) {
        XLConnect::writeWorksheetToFile(f, gene.cpd.summary, sheet = 'summary')
    }
})


```




## jointpa

```{r}
mSet<-InitDataObjects("conc", "pathinteg", FALSE)

mSet<-SetOrganism(mSet, "hsa")

geneListFile<-"replace_with_your_file_name"

geneList<-readChar(geneListFile, file.info(geneListFile)$size)

mSet<-PerformIntegGeneMapping(mSet, geneList, "hsa", "symbol");

cmpdListFile<-"replace_with_your_file_name"

cmpdList<-readChar(cmpdListFile, file.info(cmpdListFile)$size)

mSet<-PerformIntegCmpdMapping(mSet, cmpdList, "hsa", "kegg");

mSet<-CreateMappingResultTable(mSet)

mSet<-PrepareIntegData(mSet);

mSet<-PerformIntegPathwayAnalysis(mSet, "dc", "hyper", "integ", "query");

mSet<-PlotPathSummary(mSet, F, "path_view_0_", "png", 72, width=NA)

mSet<-CreateIntegMatchingTable(mSet);

mSet<-SaveTransformedData(mSet)
```



