## 120-analysis

20201031
Oxysterol 删除 squalene, Desmosterol, zymosterol

20201002
Ratio的barplot + error bar + jitter
Change unit to MFT

```{r setup, include=F, echo=F, message=F, warnings=F}
source('Rscript/setup.R')
source('Rscript/functions.R')
source('Rscript/app_functions.R')
# source('rscript/pathway_funs.R', encoding = 'UTF-8')


options(useFancyQuotes = FALSE)
options(java.parameters = '-Xmx8000m')

knitr::opts_chunk$set(
  cache = T,
  message = F,
  dpi = 300
)

library(plyr)
library(ggplot2)
library(magrittr)
# library(MetaboAnalystR)
library(kableExtra)
library(dplyr)
library(patchwork)

# library(extrafont)
# loadfonts(device = 'win', quiet = T)
# par(family = 'Liberation')
# windowsFonts(
#   serif = 'Liberation Serif',
#   sans = 'Liberation Sans',
#   mono = 'Liberation Mono'
# )

par(family = 'sans')
font_family <- 'sans'

set.seed(1)

prefix <- 'plot_2'
output.dir <- file.path('output', prefix)
# Sys.setlocale("LC_ALL","Chinese")

if (!dir.exists(output.dir)) {
    dir.create(output.dir, recursive = T)
}
```

```{r}
file10 <- 'data/heatmap_data/20201027 exosome_analytics_MFTC_Bowen.xlsx'

d10 <- local({
  name.rm <- c('SQUALENE', 'desmosterol', 'zymosterol')
  XLConnect::readWorksheetFromFile(
    file10, sheet = 'Numeric',
    check.names = F, rownames = 1
  ) %>% dplyr::select(!dplyr::all_of(name.rm))
  })



ds10 <- XLConnect::readWorksheetFromFile(
  file10, sheet = 'Group',
  check.names = F, rownames = 1
)

ds10 <- ds10[rownames(d10), ]

dv10 <- XLConnect::readWorksheetFromFile(
  file10, sheet = 'NumericGroup',
  check.names = F, rownames = 1
)

d11 <- d10[ds10$GroupID %in% c('S1', 'S2', 'S3', 'S4'), ]

ds11 <- subset(ds10, GroupID %in% c('S1', 'S2', 'S3', 'S4')) %>%
  mutate(
    GroupID = factor(GroupID)
  )

testthat::expect_equal(
  rownames(d10),
  rownames(ds10)
)

testthat::expect_equal(
  colnames(d10),
  rownames(dv10)
)

attr10 <- getAttrs(colnames(d10))

attr10[256, 'class'] <- 'CS'
attr10[574:600, 'class'] <- 'Sterol'
attr10[601:625, 'class'] <- 'Class'
attr10[626:629, 'class'] <- 'ratio'

attr10 <- attr10 %>%
  mutate(
    class2 = forcats::fct_lump_min(class, 2)
  )
```

```{r}
ht.res <- hypothesisTest(d11, ds11$GroupID)

ht.res.fdr <- ht.res %>% 
  as.data.frame() %>%
  dplyr::select(
    dplyr::starts_with(c('parametric', 'TukeyHSD', 'Games', 'non-parametric', 'Dunn'))
  ) %>%
  lapply(function(x) {p.adjust(x, method = 'fdr')}) %>%
  do.call(cbind, .)

ht.res.letters <- sapply(
  1:nrow(ht.res),
  function(i) {
    p_matrix <- matrix(1, nrow = 4, ncol = 4)
    p_matrix[lower.tri(p_matrix)] <- ht.res[i, startsWith(colnames(ht.res), 'Dunn')]
    rownames(p_matrix) <- colnames(p_matrix) <- c('S1', 'S2', 'S3', 'S4')
    multcompView::multcompLetters(p_matrix)$Letters
  }) %>% t %>%
  `rownames<-`(rownames(ht.res))

# file.excel <- file.path(
#   output.dir,
#   'hypothesis_test.xlsx'
# )
# 
# if(!file.exists(file.excel)) {
#   XLConnect::writeWorksheetToFile(
#     file.excel,
#     data = ht.res,
#     sheet = 'hypothesis_test',
#     colnames = rownames(ht.res)
#   )
# }

if (F) {
  XLConnect::writeWorksheetToFile(
    file.path(output.dir,  'hypothesis_test.xlsx'),
    ht.res,
    sheet = 'P',
    rownames = 'term'
  )
  
  XLConnect::writeWorksheetToFile(
    file.path(output.dir,  'hypothesis_test.xlsx'),
    ht.res.fdr %>%
      `rownames<-`(rownames(ht.res)),
    sheet = 'FDR',
    rownames =  'term'
  )
}

```

## Heatmap

Kruskal.Wallis and Dunn's test

```{r, eval=F}
d11a <- lapply(d11, function(x) {
  vaggregate(scale(x), ds11$GroupID, mean)
}) %>% do.call(rbind, .) %>%
  `colnames<-`(c('S1', 'S2', 'S3', 'S4'))

file.pdf <- file.path(output.dir, 'Heatmap_sterol_class.pdf')


p_list <- lapply(c('Sterol', 'Class'), function(class.i) {
  print(class.i)
  
  ind <- attr10$class2 == class.i
  
  d1 <- d11a[ind, ]

  rownames(d1) <- stringr::str_pad(
    rownames(d1), max(nchar(rownames(d1))), side = 'left'
  )
  rownames(d1) <- paste0(rownames(d1),
                         ifelse(ht.res[ind, 'non-parametric pvalue'] < 0.05, ' *', '  '))
  
  d2 <- ht.res.letters[ind, ] 
  
  rownames(d2) <- rownames(d1)
  d1 <- reshape2::melt(d1)
  d2 <- reshape2::melt(d2)
  
  p <- ggplot(d1, aes(x = Var2, y = reorder(Var1, desc(Var1)))) +
    geom_tile(aes(fill = value)) +
    geom_text(data = d2, aes(x = Var2, y = Var1, label = value)) +
    scale_fill_gradient2(low = 'blue', mid = 'white', high = 'red', midpoint = 0) +
    scale_x_discrete(expand = expansion(add = c(0.1, 0.6))) +
    labs(x = '', y = '', title = '') +
    guides(fill = guide_colorbar(title = NULL, barheight = unit(0.8, 'npc'))) +
    theme_minimal(base_family = font_family) +
    theme(panel.grid.major = element_blank(),
          axis.text.x = element_text(size = 12),
          plot.title = element_text(hjust = 0.5))
  
  p
})

if (!file.exists(file.pdf)) {
  cairo_pdf(file.pdf, onefile = T, width = 10, height = 6, family = font_family)
  
  p_list[[1]] + p_list[[2]]
  
  dev.off()
}

```

## Volcano plot

```{r, eval=F}
source('rscript/app_functions.R')

file.pdf <- file.path(output.dir, 'volcano.pdf')

g_combs <- combn(c('S1', 'S2', 'S3', 'S4'), 2)

cairo_pdf(file.pdf, onefile = T, width = 6, height = 5, family = font_family)

plyr::l_ply(1:ncol(g_combs), function(i) {
  g1 <- g_combs[1, i]
  g2 <- g_combs[2, i]

  plyr::l_ply(c('Sterol', 'lipid'), function(class.i) {
    if (class.i == 'Sterol') {
      ind <- which(attr10$class2 == class.i)
      
    } else {
      ind  <-  which(!(attr10$class2 %in% c('Sterol', 'ratio')))
    }
    
    d1 <- data.frame(
      p = ht.res[ind, sprintf('Dunn: %s:%s', g2, g1)],
      fold = ht.res[ind, sprintf('Fold: %s/%s', g2, g1)],
      label = rownames(ht.res)[ind]
    )
    
    p1 <- plot_volcano(d1, fold_cutoff = 1.2, max_label = 10,
                 title = paste0(g2, '/', g1)) 
    
    print(p1)
    
    d1 <- data.frame(
      p = ht.res.fdr[ind, sprintf('Dunn: %s:%s', g2, g1)],
      fold = ht.res[ind, sprintf('Fold: %s/%s', g2, g1)],
      label = rownames(ht.res)[ind]
    )
    
    p2 <- plot_volcano(d1, fold_cutoff = 1.2, max_label = 10,
                 title = paste0(g2, '/', g1)) +
      labs(y = '-log10(FDR)') 

    print(p2)
  })
})

dev.off()
```


```{r}
plot_volcano <- function(d, sig_lvl = 0.05, fold_cutoff = 1.5, 
                         max_label = 10, title = '', 
                         ggtheme = theme_bw(), font_family = 'Arial',
                         force = 1, expand = 0.05,
                         min.segment.length = 0.5) {
    ## d is data.frame with 3 columns: pvalue, fold, label
    ## sig_lvl is the significance level
    ## fold_cutoff is fold cutoff
    assertthat::assert_that(ncol(d) == 3)
    
    colnames(d) <- c('pvalue', 'fold', 'label')
    
    d <- d %>% dplyr::mutate(
        # pvalue and fold change both must pass threshold
        significant = ifelse(
            pvalue >= sig_lvl, 'Not significant',
            ifelse(
                fold > fold_cutoff, 
                paste0('P<', sig_lvl, '&FC>', fold_cutoff),
                ifelse(
                    fold < 1/fold_cutoff,
                    paste0('P<', sig_lvl, '&FC<1/', fold_cutoff),
                    'Not significant'
                ))),
        label = ifelse(significant != 'Not significant', label, '')
    ) %>% dplyr::filter(
        !is.na(pvalue) & !is.na(fold) & is.finite(fold) & fold > 0
    )
    
    
    if (max_label > 0) {
        ## number of significant variables
        n_sig <- length(which(d$label != ''))
        if (n_sig > max_label) {
            ## ids of top max_label smallest P value
            ## order by P value followed by fold change
            id_label <- intersect(
                # order(rank(d[, 'pvalue']) + rank(-abs(log2(d[, 'fold'])))),
                order(d[, 'pvalue']),
                which(d$significant != 'Not significant')
            )[1:max_label]
            d[-id_label, 'label'] <- ''
        }
    } else if(max_label == 0) {
        d[, 'label'] <- ''
    }
    
    if (fold_cutoff == 1) {
        ## change significant to P > 0.05 and P < 0.05
        ## change color_values accordingly
        d <- d %>% dplyr::mutate(
            significant = ifelse(
                significant == 'Not significant',
                paste0('P\u2265', sig_lvl),
                paste0('P<', sig_lvl)
            )
        )
        color_values <- c('gray40', 'red') %>%
            `names<-`(c(paste0('P\u2265', sig_lvl),
                        paste0('P<', sig_lvl)))
    } else {
        color_values <- c('gray40', 'red', 'green') %>%
            `names<-`(c('Not significant', 
                        paste0('P<', sig_lvl, '&FC>', fold_cutoff),
                        paste0('P<', sig_lvl, '&FC<1/', fold_cutoff)))
    }
    
    p <- ggplot(d, aes(x = log2(fold), y = -log10(pvalue))) +
        geom_point(aes(color = significant), size = 1) +
        scale_color_manual(values = color_values) +
        ggrepel::geom_text_repel(aes(label = label), family = font_family,
                                 size = 2.5,
                                 force = force, max.iter = 2000) +
        geom_vline(xintercept = 0, linetype = 'dashed', alpha = 0.5) +
        geom_hline(yintercept = -log10(sig_lvl), linetype = 'dashed', alpha = 0.5) +
        scale_x_continuous(
            limits = c(-max(abs(log2(d$fold))), max(abs(log2(d$fold)))),
            expand = expansion(mult = c(expand, expand), add = c(0, 0))) +
        scale_y_continuous(expand = expansion(mult = c(0, expand), add = c(0, 0))) +
        guides(color = guide_legend(title = 'Significant')) +
        labs(title = title, x = 'log2(fold change)', y = '-log10(P value)') +
        ggtheme
    
    ## add vertical corresponds to fold change cut-off
    if (fold_cutoff > 1) {
        p <- p +
            geom_vline(xintercept = log2(fold_cutoff), linetype = 'dashed', alpha = 0.5) +
            geom_vline(xintercept = -log2(fold_cutoff), linetype = 'dashed', alpha = 0.5)
    }
    
    p
}

file.pdf <- file.path(output.dir, 'volcano_lipids.pdf')

g_combs <- combn(c('S1', 'S2', 'S3', 'S4'), 2)[, c(1, 3)]

cairo_pdf(file.pdf, onefile = T, width = 9, height = 4, family = font_family)

p_list <- lapply(1:ncol(g_combs), function(i) {
  g1 <- g_combs[1, i]
  g2 <- g_combs[2, i]

  class.i <- 'lipids'
  
  if (class.i == 'Sterol') {
    ind <- which(attr10$class2 == class.i)
    
  } else {
    ind  <-  which(!(attr10$class2 %in% c('Sterol', 'ratio')))
  }
  
  d1 <- data.frame(
    p = ht.res[ind, sprintf('Dunn: %s:%s', g2, g1)],
    fold = ht.res[ind, sprintf('Fold: %s/%s', g2, g1)],
    label = rownames(ht.res)[ind]
  )
  
  # update_geom_defaults('text', list(font_size = update_geom_defaults1))
  p1 <- plot_volcano(d1, fold_cutoff = 1.2, max_label = 10, force = 10,
                     title = paste0(g2, '/', g1)) 
})

p_list[[1]] + p_list[[2]] +
  plot_layout(guides = 'collect')

dev.off()



file.pdf <- file.path(output.dir, 'volcano_sterol.pdf')

g_combs <- combn(c('S1', 'S2', 'S3', 'S4'), 2)[, c(1, 3)]

cairo_pdf(file.pdf, onefile = T, width = 9, height = 4, family = font_family)

p_list <- lapply(1:ncol(g_combs), function(i) {
  g1 <- g_combs[1, i]
  g2 <- g_combs[2, i]

  class.i <- 'Sterol'
  
  if (class.i == 'Sterol') {
    ind <- which(attr10$class2 == class.i)
    
  } else {
    ind  <-  which(!(attr10$class2 %in% c('Sterol', 'ratio')))
  }
  
  d1 <- data.frame(
    p = ht.res[ind, sprintf('Dunn: %s:%s', g2, g1)],
    fold = ht.res[ind, sprintf('Fold: %s/%s', g2, g1)],
    label = rownames(ht.res)[ind]
  )
  
  # update_geom_defaults('text', list(font_size = update_geom_defaults1))
  p1 <- plot_volcano(d1, fold_cutoff = 1.2, max_label = 20, force = 10,
                     title = paste0(g2, '/', g1)) 
})

p_list[[1]] + p_list[[2]] +
  plot_layout(guides = 'collect')

dev.off()
```

## Heatmap of fold change

P < 0.05

```{r}
file.pdf <- file.path(output.dir, 'heatmap_fold_change_fdr.pdf')

if (!file.exists(file.pdf)) {
  cairo_pdf(file.pdf, onefile = T, width = 5, height = 8, family = font_family)
  
  d.fold <- ht.res[1:625, startsWith(colnames(ht.res), 'Fold')]
  d.fdr <- ht.res[1:625, startsWith(colnames(ht.res), 'Dunn')] %>%
    apply(2, p.adjust, method = 'fdr') < 0.05
  # rownames(d.fold) <- stringr::str_trunc(rownames(d.fold), 40)
  
  d.fold[!d.fdr] <- 1
  
  ind <- which(rowSums(d.fdr) > 0)
  ymin <- min(log2(d.fold[ind, ]))
  ymax <- max(log2(d.fold[ind, ]))
  
  
  pheatmap::pheatmap(log2(d.fold[ind, ]), 
                     main = '',
                     breaks = c(-Inf, 
                                seq(ymin, -0.0001, length.out = 49),
                                0,
                                seq(0.0001, ymax, length.out = 49),
                                Inf),
                     color = colorRampPalette(c('navyblue', 'white', 'firebrick'))(100),
                     cluster_rows = T, cluster_cols = F,
                     border = 'white',
                     clustering_method = 'ward.D2',
                     na_col = 'white')
  
  dev.off()
}

file.pdf <- file.path(output.dir, 'heatmap_fold_change_p.pdf')

if (!file.exists(file.pdf)) {
  cairo_pdf(file.pdf, onefile = T, width = 5, height = 20, family = font_family)
  
  d.fold <- ht.res[1:625, startsWith(colnames(ht.res), 'Fold')]
  d.fdr <- ht.res[1:625, startsWith(colnames(ht.res), 'Dunn')] < 0.05
  # rownames(d.fold) <- stringr::str_trunc(rownames(d.fold), 40)
  
  d.fold[!d.fdr] <- 1
  
  ind <- which(rowSums(d.fdr) > 0)
  ymin <- min(log2(d.fold[ind, ]))
  ymax <- max(log2(d.fold[ind, ]))
  
  
  pheatmap::pheatmap(log2(d.fold[ind, ]), 
                     main = '',
                     breaks = c(-Inf, 
                                seq(ymin, -0.0001, length.out = 49),
                                0,
                                seq(0.0001, ymax, length.out = 49),
                                Inf),
                     color = colorRampPalette(c('navyblue', 'white', 'firebrick'))(100),
                     cluster_rows = T, cluster_cols = F,
                     border = 'white',
                     clustering_method = 'ward.D2',
                     na_col = 'white', fontsize_row = 4)
  
  dev.off()
}


```

## point line

P value

```{r}
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}

cols <- gg_color_hue(4)

d1 <- data.frame(
  Group = ds11$GroupID,
  d11,
  check.names = F
)

file.pdf <- file.path(output.dir, 'line_timepoints.pdf')

if (!file.exists(file.pdf)) {
  cairo_pdf(file.pdf, onefile = T,
          width = 4, height = 3)
  
  plyr::l_ply(
    2:ncol(d1),
    function(i) {
      ## Wuhan
      p_matrix1 <- matrix(1, nrow = 4, ncol = 4)
      p_matrix1[lower.tri(p_matrix1)] <- ht.res[i - 1,                                                startsWith(colnames(ht.res), 'Dunn')]
      colnames(p_matrix1) <- rownames(p_matrix1) <- levels(d1$Group)
      sig_letters <- multcompView::multcompLetters(p_matrix1, threshold = 0.05)
      df1 <- data.frame(
        Group  = names(sig_letters$Letters),
        sig_letters = sig_letters$Letters)
      
      y_max <- max(d1[, i])
            
      p <- ggplot(d1, aes(x = Group, y = eval(parse(text = paste0('`', colnames(d1)[i], '`'))))) +
        geom_jitter(aes(color = Group), alpha = 0.5,
                    position = position_jitter(width = 0.2)) +
        scale_color_manual(values = cols) +
        stat_summary(aes(color = Group),
                     fun.data = mean_cl_normal, fun.args = list(mult = 1),
                     geom = 'errorbar', width = 0.5) +
        stat_summary(aes(color = Group), fun = mean, 
                     geom = 'point', size = 3,
                     position = position_dodge(width = 0.5)) +
        stat_summary(aes(group = 1), fun = mean, 
                     geom = 'line') +
        geom_text(aes(x = Group, y = 1.15*y_max, label = sig_letters),
                  data = df1, color = cols) +
        labs(x = '', y = 'Abundance (MFTC)') +
        theme_classic()
        
      print(p)

    })
  
  dev.off()
}

```


FDR

```{r}
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}

cols <- gg_color_hue(4)

d1 <- data.frame(
  Group = ds11$GroupID,
  d11,
  check.names = F
)

file.pdf <- file.path(output.dir, 'line_timepoints_FDR.pdf')

if (!file.exists(file.pdf)) {
  cairo_pdf(file.pdf, onefile = T,
          width = 4, height = 3)
  
  plyr::l_ply(
    2:ncol(d1),
    function(i) {
      ## Wuhan
      p_matrix1 <- matrix(1, nrow = 4, ncol = 4)
      p_matrix1[lower.tri(p_matrix1)] <- ht.res.fdr[i - 1,                                                startsWith(colnames(ht.res.fdr), 'Dunn')]
      colnames(p_matrix1) <- rownames(p_matrix1) <- levels(d1$Group)
      sig_letters <- multcompView::multcompLetters(p_matrix1, threshold = 0.05)
      df1 <- data.frame(
        Group  = names(sig_letters$Letters),
        sig_letters = sig_letters$Letters)
      
      y_max <- max(d1[, i])
            
      p <- ggplot(d1, aes(x = Group, y = eval(parse(text = paste0('`', colnames(d1)[i], '`'))))) +
        geom_jitter(aes(color = Group), alpha = 0.5,
                    position = position_jitter(width = 0.2)) +
        scale_color_manual(values = cols) +
        stat_summary(aes(color = Group),
                     fun.data = mean_cl_normal, fun.args = list(mult = 1),
                     geom = 'errorbar', width = 0.5) +
        stat_summary(aes(color = Group), fun = mean, 
                     geom = 'point', size = 3,
                     position = position_dodge(width = 0.5)) +
        stat_summary(aes(group = 1), fun = mean, 
                     geom = 'line') +
        geom_text(aes(x = Group, y = 1.15*y_max, label = sig_letters),
                  data = df1, color = cols) +
        labs(x = '', y = 'Abundance (MFTC)') +
        theme_classic()
        
      print(p)

    })
  
  dev.off()
}

```


lineplot

```{r}
lst.block <- list(
  c('18:2-carnitine',
    '6K-5aOH-CHO',
    '16:0-carnitine',
    '18:1-carnitine',
    'acylcarnitine',
    'Gb3 d18:1/22:0',
    'Gb3 d18:1/24:1',
    'LacCer d18:1/24:1',
    'Gb3 d18:1/23:0',
    'Gb3 d18:1/24:0'),
  c('PE38:6p(16:0/22:6)',
    'PE40:6p',
    'PE38:6p',
    'PE40:6p(18:0/22:6)'),
  c('SL d18:1/22:0h',
    'SL d18:1/18:0h',
    'SL d18:1/20:0h',
    'SL d18:1/24:0h',
    'SL d18:1/22:1h',
    'SL',
    'SL d18:1/18:1h',
    'SL d18:1/20:1h'),
  c('PC42:7(22:6/20:1)',
    'PC42:2',
    'PC34:2(16:1/18:1)',
    'PC40:7(22:6/18:1)',
    'PG34:1',
    'PC34:3(16:1/18:2)',
    'PC42:4',
    'PC38:1',
    'PC38:2p',
    'PC36:3p(16:0/20:3)',
    'PC38:5(22:5/16:0)',
    'PC36:5',
    'BMP',
    'PC40:1',
    'PC34:3',
    'PC36:3',
    'PC42:6',
    'PC36:3(18:1/18:2)',
    'PC36:2',
    'PC',
    'PC40:5p(18:0/22:5)',
    'PC42:7',
    'BMP36:3(18:2/18:1)',
    'PC40:7'),
  c('PS 40:5',
    'PS 38:3',
    'LacCer d18:1/15:0',
    '24S,25-EPOXY-CHO',
    'PS 36:1',
    'lanosterol',
    'PS 36:1(18:0/18:1)',
    'PS 36:2',
    'FFA18:0',
    'PE40:6(20:2/20:4)',
    'FFA16:0',
    'FFA',
    'TAG54:3(18:1)',
    'Cer d18:0/18:0',
    'Cer d18:0/21:0',
    'Cer d18:1/24:1',
    'FFA22:6',
    'PS'),
  c('dihydrolanosterol',
    '7a-OH-cholestenone',
    'Cholesta-46-dien-3-one',
    'PS 38:4',
    'dihydro-T-MAS',
    '5a,6a-EPOXY-CHO',
    '7K-CHO')
)

file.pdf <- file.path(output.dir, 'line.pdf')

if (!file.exists(file.pdf)) {
  cairo_pdf(file.pdf, onefile = T, w = 4, h = 3, family = font_family)
  
  plyr::l_ply(lst.block, function(x) {
    print(x)
    d1 <- data.frame(Group = ds11$GroupID, scale(d11[, x]), check.names = F) %>%
      reshape2::melt(id.var = 'Group') %>%
      dplyr::mutate(
        variable = paste0(as.numeric(variable), ' - ', variable)
      )
    
    d2 <- d11[, x] %>% 
      sapply(function(x) {
        vaggregate(scale(x), ds11$GroupID, mean)
      }) %>%
      `rownames<-`(levels(ds11$GroupID)) %>%
      reshape2::melt() %>%
      dplyr::filter(Var1 %in% c('S1', 'S4')) %>%
      dplyr::mutate(
        Var2 = as.numeric(Var2)
      )
    
    
    p <- ggplot(d1, aes(x = Group, y = value)) +
      stat_summary(aes(group = variable, color = variable), fun = mean, 
                   geom = 'line') +
      # ggrepel::geom_text_repel(data = subset(d2, Var1 == 'S1'), nudge_x = -0.2,
      #                          force = 2,
      #                          aes(x = Var1, y = value, label = Var2)) +
      # ggrepel::geom_text_repel(data = subset(d2, Var1 == 'S4'), nudge_x = 0.2,
      #                          force = 2,
      #                          aes(x = Var1, y = value, label = Var2)) +
      labs(x = '', y = 'Normalized abundance (SD)') +
      guides(color = guide_legend(title = NULL)) +
      theme_classic() +
      theme(legend.text = element_text(size = 8),
            legend.position = 'none')
    
    print(p)
  })
  
  dev.off()
}

```


## Ratio barplot + error bar + jitter

```{r}
d1 <- data.frame(Group = ds11$GroupID,
                 d11[, c("Cho/PC-1", "SM+Cho/PC-1", "Cho/PC-2", "SM+Cho/PC-2")],
                 check.names = F) %>%
  reshape2::melt(id.var = 'Group')

d2 <- ht.res.letters[c("Cho/PC-1", "SM+Cho/PC-1", "Cho/PC-2", "SM+Cho/PC-2"), ] %>%
  reshape2::melt()

ymax <- max(d1$value)

p <- ggplot(d1, aes(x = variable, y = value)) +
  stat_summary(aes(fill = Group), geom = 'bar', fun = mean,
               alpha = 0.6,
               position = position_dodge2(width = 1)) +
  stat_summary(aes(color = Group), geom = 'errorbar',
               position = position_dodge2(width = 1),
               fun.data = mean_cl_normal, fun.args = list(mult = 1)) +
  geom_jitter(aes(color = Group), alpha = 0.5,
              position = position_jitterdodge(dodge.width = 0.9)) +
  geom_text(data = d2, aes(x = Var1, y = 1.05 * ymax, color = Var2, label = value),
            position = position_dodge2(width = 0.9), hjust = 0.5) +
  labs(y = 'Ratio', x = '') +
  theme_classic()

file.pdf <- file.path(output.dir, 'ratio_barplot.pdf')

if (!file.exists(file.pdf)) {
  cairo_pdf(file.pdf, onefile = T, width = 5, height = 3)
  print(p)
  dev.off()
}
```

## Barplot

```{r}
attr11 <- getAttrs(colnames(d11))
attr11[574:600, 'class'] <- 'Sterol & oxysterol'
attr11[601:625, 'class'] <- 'Lipid class'
attr11[626:629, 'class'] <- 'Ratio'
attr11[256, 'class'] <- 'CS'

attr11 <- attr11 %>%
  mutate(
    class = forcats::fct_lump_min(class, 2, other_level = 'Others')
  )

d3 <- lapply(d11, function(x) {
      as.data.frame(vaggregate(x, ds11$GroupID, mean_cl_normal, mult = 1))['ymax', ] %>%
      unlist
    }) %>% 
      do.call(rbind, .) %>%
      `colnames<-`(c('S1', 'S2', 'S3', 'S4')) %>%
    reshape2::melt() %>%
  magrittr::set_colnames(c('rn', 'variable', 'y'))



d2 <- data.frame(ht.res.letters, class = attr11[, 'class'], rn = rownames(ht.res.letters)) %>%
 reshape2::melt(id.vars = c('class', 'rn')) %>%
 dplyr::left_join(d3, by = c('rn', 'variable'))

 plyr::l_ply(unique(attr11[, 'class']), function(class.i) {
  ind <- which(attr11[, 'class'] == class.i)
  print(class.i)
  if (T) {
    n_page <- ceiling(length(ind) / 16)
    
    n <- length(ind)
    n_count <- 0
    
    for (i in 1:n_page) {
      ## number of species to plot
      n_cur <- ceiling((n - n_count) / (n_page - i + 1))
      
      ind_cur <- ind[(n_count + 1):(n_count + n_cur)]
      
      n_count <- n_count + n_cur
      
      ymax.class <- max(d11[, ind_cur])
      
      d1 <- data.frame(Group = ds11$GroupID,
                       d11[, ind_cur],
                       check.names = F) %>%
        reshape2::melt(id.var = 'Group')
      
      dd2 <- subset(d2, rn %in% colnames(d11)[ind_cur])
      
      p <- ggplot(d1, aes(x = variable, y = value)) +
        stat_summary(aes(fill = Group), geom = 'bar', color = 'black', 
                     fun = mean,
                     alpha = 1,
                     position = position_dodge2(width = 1)) +
        stat_summary(aes(group = Group), geom = 'errorbar', 
                     position = position_dodge2(width = 1),
                     fun.data = mean_cl_normal, fun.args = list(mult = 1)) +
        # geom_jitter(aes(color = Group), alpha = 0.5,
        #             position = position_jitterdodge(dodge.width = 0.9)) +
        geom_text(data = dd2, aes(x = rn, y = y + 0.01 * ymax.class, 
                                  label = value, color = variable),
                  size = 2.5, vjust = 0,
                  position = position_dodge2(width = 0.9)) +
        # scale_y_sqrt() +
        labs(y = 'Abundance (MFT)', x = '', title = class.i) +
        theme_classic() +
        theme(legend.position = 'none',
              plot.margin = margin(0.5, 0.5, 0, 1, unit = 'cm'),
              axis.text.x = element_text(angle = 15, hjust = 0.9, vjust = 0.9))
       
      w.pdf <- 7 / 16 * n_cur + 1

      file.pdf <- file.path(output.dir, 'barplot', paste0(make.names(class.i), '_', sprintf('%.2d', i), '.pdf'))
      file.tiff <- file.path(output.dir, 'barplot', paste0(make.names(class.i), '_', sprintf('%.2d', i), '.tiff'))

      dir.create(file.path(dirname(file.pdf)))

      
      ggsave(file.pdf, p, width = w.pdf, height = 3)
      ggsave(file.tiff, p, width = w.pdf, height = 3,
             units = 'in', compression = 'lzw+p', dpi = 300)
    }
  }
  
})

```

